Experiments and Modeling of the Autoignition of Methyl Pentanoate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine

### 2017-09-27 revision-1/response/response.tex

\begin{reviewer}
%
    Please improving the quality of figures. For eg: First stage and total ignition delay Symbols
    in Figs 5a \& 5b are very hard to differentiate. Figure 1 could also be improved.
%
\end{reviewer}

\begin{response}
    %
    We have increased the mark size on \cref{fig:simulation-comparison} to better distinguish the
    overall ignition delay and the first-stage ignition delay. On \cref{fig:ign-delay-def}, we made
    several changes to improve the clarity, including swapping the colors of reactive and
    non-reactive pressure traces, changing the non-reactive pressure to a dashed line style, and
    making the black lines thinner.
    %
\end{response}

\begin{reviewer}
    %
    Regarding the statement "The NTC region of MV is mapped out to provide additional information on
    the fidelity of using MV as a biodiesel surrogate." I am not convinced that MV could be a
    biodiesel surrogate, in my opinion it helps in understanding the chemistry of large esters.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that it is possible that MV will not be a good surrogate for
    biodiesel. Nonetheless, to our knowledge, there has been no study to suggest that it is not a
    good surrogate, probably because very few studies of MV have been conducted to date, so there is
    not enough information to judge. However, we think that judging the suitability (or lack
    thereof) for MV as a surrogate of biodiesel is outside the scope of this work, so we are simply
    noting that this study provides part of the necessary experimental information so that others
    may judge for themselves. We also agree with the reviewer that the scope of this work is
    actually wider than we had noted previously, so we have added a note that this study provides
    further information on the autoignition chemistry of large methyl esters near line 41:

@@

\begin{reviewer}
    Specify the units of coefficients B, C  and, temperature T in equation (1)
\end{reviewer}

\begin{response}
    The coefficients are in \si{kPa} and \si{\kelvin}. This has been added to the text near line
    112 and in the caption of \cref{tab:antoine}.
    \begin{quote}
        ``... where \(A\), \(B\), and \(C\) are substance-specific coefficients, given in units of
        \si{\kelvin} and \si{\kPa}.''
    \end{quote}
\end{response}

\begin{reviewer}
    %
    The authors report an ignition delay of 100 ms at 640 K f=2, MV/air mixture at 15 bar. So at 30
    bar I would expect an ignition delay which is for sure will be more than 10 ms. So I am
    surprised by the statement "For the phi = 2:0 condition, only  PC = 15 bar is considered because
    we could not achieve TC values low enough that the ignition during the compression stroke can be
    prevented'. Can the authors comment on this.
    %
\end{reviewer}

\begin{response}
    %
    The reason for this is because of physical limitations of the experimental apparatus and the
    fuel. In particular, achieving sufficiently EOC low temperatures on our apparatus would require
    increasing the reaction chamber volume at the end of compression, which in turn requires
    substantially higher initial pressures to reach the EOC pressure of \SI{30}{\bar}. The initial
    pressure in the RCM reaction chamber is limited by the pressure in the mixing tank, and the
    total pressure in the mixing tank is limited by the vapor pressure of MV at the preheat
    temperature. The combination of these factors meant that we could not prepare a mixture with
    sufficiently high tank pressure to enable repeated experiments with the EOC volume required to
    reach EOC temperatures where the ignition delay would be measurable. However, we feel that this
    explanation would only confuse the manuscript, so we omit it.
    %
\end{response}

@@

\begin{reviewer}
    %
    The comparison of ignition delays of MV from current work and that of Hadj-Ali et al .[9] would
    be helpful.
    %
\end{reviewer}

\begin{response}
    %
    The ignition delays of MV measured in the work of \citet{Hadj-Ali2009} were conducted at several
    pressures and roughly a single EOC temperature (near \SI{815}{\K}). However, none of the
    pressures considered in that study match the pressures considered in our study, so we feel that
    including the data on one of our existing plots would be very confusing. Moreover, the influence
    of facility-specific effects make a direct comparison of ignition delays from different RCMs
    somewhat challenging, particularly without experimental pressure traces of MV from the work of
    \citet{Hadj-Ali2009} to compare the facility effects. Therefore, we have not added a figure with
    this comparison.
    %
\end{response}

\begin{reviewer}
    %
    On discussion related to Dievart model in figure 5, it is surprising that phi=0.25 and phi=0.5
    simulations show NTC but phi=1 and phi=2 simulations do not. The trend is counter intuitive as
    the cool flame chemistry and NTC response are expected to increase with increase in equivalence
    ratio. This needs to be clarified. My guess is that the location of the NTC is not captured
    accurately by model. NTC region predicted by model is shifted to high temperatures where
    ignition is observed in RCM simulations. Showing Adiabatic constant volume simulations to
    understand the location of NTC at different equivalence ratios.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that the location of the NTC region with respect to inverse
    temperature is not well captured by the model. It does seem as though the NTC has been shifted
    to higher temperatures with respect to the data, in addition to the model being too reactive.
    This is supported by the newly added \cref{fig:conv-comparison}, which compares adiabatic,
    constant volume simulations with the RCM simulations from \cref{fig:simulation-comparison}. It
    is apparent from this figure that the predicted ignition delays in the $P_C = \SI{15}{\bar}$,
    $\phi = \num{2.0}$ and the $P_C = \SI{30}{\bar}$, $\phi = \num{1.0}$ simulations fall on the low
    temperature side of the NTC, while the $P_C = \SI{15}{\bar}$, $\phi =$ \numlist{0.25;0.50}
    simulated ignition delays are on the high-temperature side of the NTC, but approaching the NTC
    range such that they start to assume the characteristic curvature as the temperature decreases.
    In addition to the figure, we have added some paragraphs near line 293 to clarify this point,
    but we have not copied the added text here to save space.
    %
\end{response}

@@

\begin{reviewer}
    %
    Regarding the heat of formation shown in Table 3, it is interesting that the heat of formation
    of radical 3 and 4 estimated by RMG are different. I would request the authors check their
    numbers.
    %
\end{reviewer}

\begin{response}
    %
    We have checked the values for the estimates in the mechanism, and they are identical to the
    values in \cref{tab:heat-of-formation} and further identical to the values available from the
    RMG website at \url{http://rmg.mit.edu}. As such, no changes have been made to the manuscript in
    response to this comment.
    %
\end{response}

\begin{reviewer}
    %
    The authors use the branching ratios of fuel decomposition to fuel radicals to state the
    importance of fuel chemistry in the current work. The relative consumption shown in Table 4
    demonstrates that the branching ratios do not exhibit strong sensitive to thermochemistry which
    is supported by the fact that RMG and RMG switched show nearly identical branching ratios. I am
    not convinced that the importance of the thermochemistry is elucidated by the current discussion
    related to Table 4 and the discussion in my opinion does not provide any kinetic insights. May
    be the discussion could be removed.
    %
\end{reviewer}

\begin{response}
    %
    We have clarified the discussion of this point. In particular, we have emphasized that the
    results show that both the thermochemistry and reaction pathways are important, and cannot
    easily be disentangled from each other, requiring a thorough, detailed investigation of the
    methyl valerate reaction system which we consider outside the scope of the present work. In
    addition, we have noted that improper prediction of the thermochemistry of species may affect
    the RMG-generated model in particular due to the rate-based algorithm used in RMG, which may
    miss important reactions if the thermochemistry is not accurate. We have modified the text
    beginning near line 402, but do not copy the modification here for space.
    %
\end{response}

@@

\begin{reviewer}
    %
    Alternatively, it would be interesting to compare the thermochemistry of the low temperature
    chemistry related reactions (RO2 chemistry) from models and ab-initio works like those Hayes and
    Burgess and also try adopting the reaction rates from ab-initio calculations and see to if the
    mechanism performance improves.
    %
\end{reviewer}

\begin{response}
    %
    We feel that the poor performance of both mechanisms indicates that a thorough and detailed
    study of the entire methyl valerate system is warranted, and small modifications, such as the
    replacement of only a few rates related to low-temperature chemistry, will not be sufficient to
    significantly improve the performance of the model. Moreover, we feel that such ad-hoc
    replacement of rates, even if the performance improves, could not be shown to be anything more
    than fortuitous without further detailed study, which is outside the scope of the present work.
    Finally, in a separate work, Di\'{e}vart et al. (Combustion and Flame, 159 (5) 1793--1805, 2012)
    note that ``preliminary tests with up-to-date kinetic parameters [including the work of Hayes
    and Burgess] for the generic low temperature scheme have shown an enhancement of the fuel
    reactivity at the end of the NTC region.'' Since the model already over-predicts the reactivity
    of the RCM data, particularly at the end of the NTC region, further enhancement is not likely to
    improve agreement.

    Therefore, we have not included these modifications in the manuscript; we have, however, added a
    note to the end of the discussion (near line 418) that references the work of \citet{Hayes2009}
    to indicate that some of the necessary detailed work has been started.
    %
\end{response}

@@

\begin{reviewer}
    %
    The measurement of Methyl Valerate at these temperature and pressure conditions constitutes
    important work and the manuscript of well laid out and has potential to be widely circulated.
    The arguments in the article are sound and the figures are clear.
    %
\end{reviewer}

\begin{response}
    Thank you for your comments. A detailed response is below.
\end{response}

\begin{reviewer}
    %
    I would like to ask the authors to check one item definitively before the article is published.
    I am concerned that at these higher pressures and equivalence ratios close to 1, there would be
    ignition which occurs during the compression stroke. This would not be noticed if some of the
    fuel condenses during the charging process and thereby you are dealing with lean mixtures than
    those cited in the paper. The authors are encouraged to make sure this is not the case by
    possibly sampling the charge after filling through a GC or by a quantitative absorption method
    as that used routinely by Stanford.
    %
\end{reviewer}

\begin{response}
    %
    With regards to ignition during the compression stroke, we have compared each reactive pressure
    trace to its corresponding non-reactive trace and found they are in good agreement from the
    beginning to the end of compression, indicating that there is no heat release during the
    compression stroke. With regards to the filling process, we have thoroughly checked the
    temperatures throughout the apparatus and found them to be quite consistent, including the
    mixing tanks, the piping connections, and the reaction chamber itself. Moreover, we have
    conducted several studies in the past utilizing GC to determine mole fractions of components.
    All of these studies determined that the vaporization procedure used on this apparatus is
    sufficient to produce a homogeneous mixture of fuel and oxidizer, even for fuels with relatively
    low vapor pressures. We have added references to these prior studies in the present manuscript
    near line 110:

### 2017-09-27 revision-1/revised-paper/methyl-valerate.tex
Final figure updates and response checks

\begin{center}
    \captionsetup{type=figure}
    \includegraphics[width=\textwidth]{figures/constant-volume-comparison.pdf}
    \caption{Comparison of simulated ignition delays computed in an adiabatic, constant volume
    system (dotted lines) and computed using the procedure described in
    \cref{sec:experimental-modeling} (solid lines). a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
    \label{fig:conv-comparison}
\end{center}

To further understand the model of \citet{Dievart2013}, we have conducted adiabatic, constant volume
simulations (called CONV simulations), as these simulations are not linked to a particular
experiment by the volume trace and can be conducted over a wide range of temperatures. In the CONV
simulations, overall ignition delay is defined as an increase in the temperature of \SI{400}{\K}
over the initial temperature. \Cref{fig:conv-comparison} compares the CONV simulations to the RCM
simulations shown in \cref{fig:simulation-comparison}. From \cref{fig:conv-comparison}, it can be
seen that the curvature in the RCM simulations at \(P_C = \SI{15}{\bar}\), \(\phi =\)
\numlist{0.25;0.5} is related to the NTC region of the overall ignition delay, while the lack of
curvature in the \(P_C = \SI{15}{\bar}\), \(\phi = \num{2.0}\) and the \(P_C = \SI{30}{\bar}\),
\(\phi = \num{1.0}\) simulations is because those lie on the low-temperature side of the predicted
NTC. It is clear from \cref{fig:conv-comparison} that the model of \citet{Dievart2013} predicts the
NTC to occur at too high a temperature relative to the experiments.

>>>>

To further understand the model of \citet{Dievart2013}, we have conducted adiabatic, constant volume
simulations (called CONV simulations), as these simulations are not linked to a particular
experiment by the volume trace and can be conducted over a wide range of temperatures. In the CONV
simulations, overall ignition delay is defined as an increase in the temperature of \SI{400}{\K}
over the initial temperature. The diluent for all of the CONV simulations is pure argon, although
the RCM simulations described previously consider the diluent mixture associated with each
experiment (either pure argon or a 50:50 mixture of argon and nitrogen by mole, see
\cref{tab:mixtures} and the Supplementary Material).

\begin{center}
    \captionsetup{type=figure}
    \includegraphics[width=\textwidth]{figures/constant-volume-comparison.pdf}
    \caption{Comparison of simulated overall ignition delays computed in an adiabatic, constant
    volume system (dotted lines) and computed using the procedure described in
    \cref{sec:experimental-modeling} (solid lines). a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
    \label{fig:conv-comparison}
\end{center}

\Cref{fig:conv-comparison} compares the CONV simulations to the RCM simulations shown in
\cref{fig:simulation-comparison}. From \cref{fig:conv-comparison}, it can be seen that the curvature
in the RCM simulations at \(P_C = \SI{15}{\bar}\), \(\phi =\) \numlist{0.25;0.5} is related to the
NTC region of the overall ignition delay, while the lack of curvature in the \(P_C =
\SI{15}{\bar}\), \(\phi = \num{2.0}\) and the \(P_C = \SI{30}{\bar}\), \(\phi = \num{1.0}\)
simulations is because those lie on the low-temperature side of the predicted NTC. It is clear from
\cref{fig:conv-comparison} that the model of \citet{Dievart2013} predicts the NTC to occur at too
high a temperature relative to the experiments.
